---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rangeR

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/rangeR)](https://CRAN.R-project.org/package=rangeR)
<!-- badges: end -->

The goal of rangeR is to provide data about the total number of people working in Protected Areas (PAs) on the planet, as well as identifying what contributes to variation in the workforce between PAs.

## Installation

You can install the development version of rangeR from [GitHub](https://github.com/) with:

``` r
# install.packages("remotes")
remotes::install_github("courtiol/rangeR")
```
## Example

This is a basic example which shows you how to solve a common problem:

```{r example, eval = FALSE}
library(rangeR)

data_rangers <- fetch_data_rangers()

if (require("skimr")) {
  skim(data_rangers)
}

## more to come
```
