---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# rangeRinPA

<!-- badges: start -->
[![Lifecycle: experimental](https://img.shields.io/badge/lifecycle-experimental-orange.svg)](https://lifecycle.r-lib.org/articles/stages.html#experimental)
[![CRAN status](https://www.r-pkg.org/badges/version/rangeR)](https://CRAN.R-project.org/package=rangeR)
<!-- badges: end -->

The goal of rangeRinPA is to reproduce most results from the paper "Protected Area Personnel and Ranger Numbers are Insufficient to Deliver Global Expectations" by Appleton et al. (submitted).

Note: this package was previously called rangeR but we changed the name since rangeRinPA is now using another package called ranger (to run random forests) with which it conflicted due to the similar name.

## Installation

### Option 1

You can install the development version of rangeRinPA from [GitHub](https://github.com/) by clicking on the green button above with the label **"Code"** and then on **Download ZIP** (do not unzip it!).

Then, in R type:

```r
# install.packages("remotes")
remotes::install_local(path = file.choose(), dependencies = TRUE, build = TRUE, build_vignettes = TRUE)
```

It will open a windows for you to select the file `rangeR-master.zip` that you just have downloaded. 


### Option 2

You can also install the development version of rangeRinPA with:

``` r
# install.packages("remotes")
remotes::install_github("courtiol/rangeR", dependencies = TRUE, build_vignettes = TRUE)
```

One difficulty however is that since this is a private GitHub repository, you may have to set up your R system so that it "knows" you GitHub credentials.

If you really want to do that, you first need to create a **Personal access tokens** in your GitHub settings (see [here for how to](https://docs.github.com/en/github/authenticating-to-github/creating-a-personal-access-token).

Then you need to add your newly created token into your .Renviron file, which you can access using `usethis::edit_r_environ()` (you need to install the package `{usethis}`, if you don't have it).

Your .Renviron file should thus have a line as the following (with a different set of letters and numbers):

```txt
GITHUB_PAT = 21b10aXXXXXXXXXXXXXXXXXXXXXdc4c276
```

## How to access the vignettes?

The different analyses we have done are stored in the package as HTML vignettes.

To view such vignettes, simply install the package (see above) and then type:

```{r, eval=FALSE}
browseVignettes("rangeRinPA")
```

Note: if the vignettes are not found, it is probably because you did not successfully install the package.
In particular, you must type `build_vignettes = TRUE` while installing the package as indicated above.

## How to use our data?

The data are embedded in the package, so you can use them after installing the package (see above) as any other data frame or tibble in R.

Here is a simple example:

```{r example}
library(rangeRinPA)

sum(data_rangers$staff_total, na.rm = TRUE)
```
