---
title: "Data completeness"
output: 
  rmarkdown::html_document:
    df_print: paged
    toc: true
    toc_depth: 3
    code_folding: hide
vignette: >
  %\VignetteIndexEntry{Data_completeness}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

<style type="text/css">
.main-container {
  max-width: 1200px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  error = TRUE,
  comment = "#>",
  cache = FALSE, # don't put TRUE as inline chunk make R CMD check to crash
  cache.path = "cache/cache_vignette_data_completeness/",
  fig.path = "figs/fig_vignette_data_completeness/",
  fig.align = 'center'
)

library(rangeR)

# set how to represent countries in plots:
#data_rangers$countryname <- countrycode(sourcevar = data_rangers$countryname_iso, "iso3c", "unicode.symbol")
#data_rangers$countryname[data_rangers$countryname_iso == "ZZZ"] <- "ZZZ"
#data_rangers$countryname[data_rangers$countryname_iso == "KOS"] <- "KOS"
data_rangers$countryname <- data_rangers$countryname_iso

# set overlap allowed in plots 
options("ggrepel.max.overlaps" = Inf)
```

```{r setup, message=FALSE}
library(dplyr)
library(tibble)
library(ggplot2)
library(forcats)
library(tidyr)
library(ggrepel)
library(spaMM)
library(countrycode)
library(gt)
```

## Rangers

Here is a plot showing the breakdown of the number of countries for which we retrieved the number of rangers:

```{r completness_rangers computation}
completness_rangers <- tibble(
  Category = c("countries for which the numbers for rangers and other staff are both separately known",
               "countries for which the numbers for rangers and other staff are both known but undistinguishable",
               "countries for which only the number for rangers is known (other staff missing)",
               "countries for which we have no about rangers or staff"),
  N = c(sum(!is.na(data_rangers$staff_rangers_others_known)),
        sum(!is.na(data_rangers$staff_total_details_unknown)),
        sum(!is.na(data_rangers$staff_rangers_others_unknown)),
        sum((is.na(data_rangers$staff_rangers_others_known) &
             is.na(data_rangers$staff_rangers_others_unknown) &
             is.na(data_rangers$staff_total_details_unknown)))
        ))
```

```{r completness_rangers pie}
ggplot(completness_rangers) +
  aes(x = "", y = N, fill = Category, label = N) +
  geom_bar(stat = "identity") +
  geom_text(position = position_stack(vjust = 0.5)) +
  coord_polar("y", start = 0) +
  theme_void() +
  theme(legend.position = "top") +
  guides(fill = guide_legend(nrow = 4, byrow = TRUE)) +
  labs(fill = "")
```

**Note**: so far the numbers don't add up due to a data issue with **Kenya**...

**In summary:** for **`r sum(is.na(data_rangers$staff_rangers))`** we have no information on rangers per se, including **`r sum(is.na(data_rangers$staff_rangers_others_known) & is.na(data_rangers$staff_rangers_others_unknown) & is.na(data_rangers$staff_total_details_unknown))`** countries for which we have no information what-so-ever.

The countries for which we miss information about rangers or staff are the following ones (ordered by decreasing country area):
```{r completness_rangers list}
data_rangers %>%
  filter(is.na(staff_rangers_others_known) &
          is.na(staff_rangers_others_unknown) &
          is.na(staff_total_details_unknown)) %>%
  select(countryname_eng) %>%
  mutate(info = "none") -> no_info_countries

data_rangers %>%
  filter(is.na(staff_rangers)) %>%
  select(countryname_iso, countryname_eng, area_country) -> partial_info_countries
  
full_join(no_info_countries, partial_info_countries, by = "countryname_eng") %>%
  mutate(info = if_else(is.na(info), "lumped with other staff", info)) %>%
  arrange(desc(area_country)) %>%
  mutate(flags = countrycode(sourcevar = countryname_iso, "iso3c", "unicode.symbol")) %>%
  mutate(Country = paste(countryname_eng, flags)) %>%
  select(Country, Info_on_rangers = info) %>%
  gt()
```

Here are the countries where the number of rangers is known but 0:

```{r no rangers}
data_rangers %>%
  filter(staff_rangers == 0) %>%
  select("countryname_eng", contains("staff"))
```

Here are the countries where the number of non rangers is known but 0:

```{r no non rangers, rows.print=25}
data_rangers %>%
  filter(staff_others == 0) %>%
  select("countryname_eng", contains("staff"))
```

### by continents

**Note**: tables sorted by PA area completeness

```{r continent abs}
data_rangers %>%
  group_by(country_UN_continent) %>%
  summarise(total_rangers_known = sum(staff_rangers, na.rm = TRUE),
            n_missing_countries = sum(is.na(staff_rangers)),
            PA_area_missing = sum(area_PA_WDPA[is.na(staff_rangers)]),
            area_country_missing = sum(area_country[is.na(staff_rangers)])) %>%
  arrange(PA_area_missing) %>%
  bind_rows(
    data_rangers %>%
      summarise(total_rangers_known = sum(staff_rangers, na.rm = TRUE),
                n_missing_countries = sum(is.na(staff_rangers)),
                PA_area_missing = sum(area_PA_WDPA[is.na(staff_rangers)]),
                area_country_missing = sum(area_country[is.na(staff_rangers)])) %>%
      mutate(country_UN_continent = "ALL"))
```

```{r continent prop}
data_rangers %>%
  group_by(country_UN_continent) %>%
  summarise(total_rangers_known = sum(staff_rangers, na.rm = TRUE),
            n_missing_countries_pct = 100 * sum(is.na(staff_rangers)) / n(),
            PA_area_missing_pct = 100 * sum(area_PA_WDPA[is.na(staff_rangers)]) / sum(area_PA_WDPA),
            area_country_missing_pct = 100 * sum(area_country[is.na(staff_rangers)]) / sum(area_country)) %>%
  arrange(PA_area_missing_pct) %>%
  bind_rows(
    data_rangers %>%
      summarise(total_rangers_known = sum(staff_rangers, na.rm = TRUE),
                n_missing_countries_pct = 100 * sum(is.na(staff_rangers)) / n(),
                PA_area_missing_pct = 100 * sum(area_PA_WDPA[is.na(staff_rangers)]) / sum(area_PA_WDPA),
                area_country_missing_pct = 100 * sum(area_country[is.na(staff_rangers)]) / sum(area_country)) %>%
      mutate(country_UN_continent = "ALL")) %>%
      mutate(across(where(is.numeric), round, digits = 2))
```

### by sub-continents

**Note**: tables sorted by PA area completeness

```{r subcontinent abs, rows.print=25}
data_rangers %>%
  group_by(country_UN_subcontinent) %>%
  summarise(total_rangers_known = sum(staff_rangers, na.rm = TRUE),
            n_missing_countries = sum(is.na(staff_rangers)),
            PA_area_missing = sum(area_PA_WDPA[is.na(staff_rangers)]),
            area_country_missing = sum(area_country[is.na(staff_rangers)])) %>%
  arrange(PA_area_missing) %>%
  bind_rows(
    data_rangers %>%
      summarise(total_rangers_known = sum(staff_rangers, na.rm = TRUE),
                n_missing_countries = sum(is.na(staff_rangers)),
                PA_area_missing = sum(area_PA_WDPA[is.na(staff_rangers)]),
                area_country_missing = sum(area_country[is.na(staff_rangers)])) %>%
      mutate(country_UN_subcontinent = "ALL"))
```

```{r subcontinent prop, rows.print=25}
data_rangers %>%
  group_by(country_UN_subcontinent) %>%
  summarise(total_rangers_known = sum(staff_rangers, na.rm = TRUE),
            n_missing_countries_pct = 100 * sum(is.na(staff_rangers)) / n(),
            PA_area_missing_pct = 100 * sum(area_PA_WDPA[is.na(staff_rangers)]) / sum(area_PA_WDPA),
            area_country_missing_pct = 100 * sum(area_country[is.na(staff_rangers)]) / sum(area_country)) %>%
  arrange(PA_area_missing_pct) %>%
  bind_rows(
    data_rangers %>%
      summarise(total_rangers_known = sum(staff_rangers, na.rm = TRUE),
                n_missing_countries_pct = 100 * sum(is.na(staff_rangers)) / n(),
                PA_area_missing_pct = 100 * sum(area_PA_WDPA[is.na(staff_rangers)]) / sum(area_PA_WDPA),
                area_country_missing_pct = 100 * sum(area_country[is.na(staff_rangers)]) / sum(area_country)) %>%
      mutate(country_UN_subcontinent = "ALL")) %>%
      mutate(across(where(is.numeric), round, digits = 2))
```

## Predictors

```{r predictors, rows.print=30}
data_rangers %>%
  select(-contains("staff")) %>%
  summarize(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), values_to = "missings_all", names_to = "variable") -> missing_pred1

data_rangers %>%
  filter(!is.na(staff_others)) %>%
  select(-contains("staff")) %>%
  summarize(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), values_to = "missings_rangers_known", names_to = "variable") -> missing_pred2

data_rangers %>%
  filter(is.na(staff_others)) %>%
  select(-contains("staff")) %>%
  summarize(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), values_to = "missings_rangers_unknown", names_to = "variable") -> missing_pred3

full_join(full_join(missing_pred1, missing_pred2), missing_pred3) %>%
  arrange(missings_all, missings_rangers_known, missings_rangers_unknown)
```

Same for countries larger than 15000 km2:

```{r predictors big countries, rows.print=30}
data_rangers %>%
  filter(area_country > 15000) -> data_rangers2

data_rangers2 %>%
  select(-contains("staff")) %>%
  summarize(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), values_to = "missings_all", names_to = "variable") -> missing_pred1

data_rangers2 %>%
  filter(!is.na(staff_others)) %>%
  select(-contains("staff")) %>%
  summarize(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), values_to = "missings_rangers_known", names_to = "variable") -> missing_pred2

data_rangers2 %>%
  filter(is.na(staff_others)) %>%
  select(-contains("staff")) %>%
  summarize(across(everything(), ~ sum(is.na(.)))) %>%
  pivot_longer(everything(), values_to = "missings_rangers_unknown", names_to = "variable") -> missing_pred3

full_join(full_join(missing_pred1, missing_pred2), missing_pred3) %>%
  arrange(missings_all, missings_rangers_known, missings_rangers_unknown)
```

