---
title: "Data correlates"
output: 
  rmarkdown::html_document:
    df_print: paged
    toc: true
    toc_depth: 3
    code_folding: hide
vignette: >
  %\VignetteIndexEntry{Data_correlates}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

<style type="text/css">
.main-container {
  max-width: 1200px;
  margin-left: auto;
  margin-right: auto;
}
</style>

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  error = TRUE,
  comment = "#>",
  cache = FALSE, # don't put TRUE as inline chunk make R CMD check to crash
  cache.path = "cache/cache_vignette_data_correlates/",
  fig.path = "figs/fig_vignette_data_correlates/",
  fig.align = 'center'
)

library(rangeR)

# set how to represent countries in plots:
#data_rangers$countryname <- countrycode(sourcevar = data_rangers$countryname_iso, "iso3c", "unicode.symbol")
#data_rangers$countryname[data_rangers$countryname_iso == "ZZZ"] <- "ZZZ"
#data_rangers$countryname[data_rangers$countryname_iso == "KOS"] <- "KOS"
data_rangers$countryname <- data_rangers$countryname_iso

# set overlap allowed in plots 
options("ggrepel.max.overlaps" = Inf)
```

```{r setup, message=FALSE}
library(dplyr)
library(tibble)
library(ggplot2)
library(forcats)
library(tidyr)
library(ggrepel)
library(spaMM)
library(countrycode)
library(gt)
```


## Number of other staff

```{r corr staff_others lm}
fit  <- lm(log(staff_rangers + 1) ~ log(staff_others + 1), data = data_rangers %>% filter(staff_rangers > 0, staff_others > 0))
coef_fit <- coef(fit)
f <- function(x) (x + 1)^coef_fit[2] * (exp(coef_fit[1]) - 1)
```

```{r corr staff_others plot, fig.height=8, fig.width=8, cache=TRUE}
data_rangers %>%
  ggplot() +
  aes(y = staff_rangers, x = staff_others, label = countryname, size = area_country, colour = country_UN_continent) +
  geom_text_repel(nudge_x = 0, nudge_y = 0, force = 0.03, force_pull = 2, seed = 123, min.segment.length = 0) +
  #stat_function(fun = f, colour = "blue", n = 1e5) +
  coord_trans(x = "log1p", y = "log1p") +
  scale_size_continuous(trans = "sqrt", guide = "none", range = c(3, 7)) +
  scale_colour_discrete(guide = "none") +
  scale_x_continuous(breaks = c(0, 10, 100, 1000, 10000), minor_breaks = NULL) +
  scale_y_continuous(breaks = c(0, 10, 100, 1000, 10000), minor_breaks = NULL) +
  theme_minimal() +
  labs(x = "Number of other staffs",
       y = "Number of rangers")
```

**Conclusion**: there is a nice (expected) relationship between the number of rangers and the the number of other staff.
Since the slope of the linear model is `r round(coef_fit[2], 2)`, the number of rangers increases less quickly than the number of other staff.


## area PA

First, let's explore the 3 variables we have:

```{r corr PA_areas plot, fig.height=8, fig.width=8, cache=TRUE}
data_rangers %>%
  ggplot() +
  aes(y = area_PA_WDPA, x = area_PA_surveyed, label = countryname, size = area_country, colour = country_UN_continent) +
  #geom_text_repel(nudge_x = 0, nudge_y = 0, force = 0.03, force_pull = 2, seed = 123, min.segment.length = 0) +
  geom_text() +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") + 
  coord_trans(x = "log1p", y = "log1p") +
  scale_size_continuous(trans = "sqrt", guide = "none", range = c(3, 7)) +
  scale_colour_discrete(guide = "none") +
  scale_x_continuous(breaks = c(0, 10, 100, 1000, 1e4, 1e5, 1e6, 1e7), minor_breaks = NULL) +
  scale_y_continuous(breaks = c(0, 10, 100, 1000, 1e4, 1e5, 1e6, 1e7), minor_breaks = NULL) +
  theme_minimal() +
  labs(x = "Area PA according to country",
       y = "Area PA according to WDPA")
```


```{r corr PA_areas plot 2, fig.height=8, fig.width=8, cache=TRUE}
data_rangers %>%
  ggplot() +
  aes(y = area_PA_WDPA, x = area_PA_total, label = countryname, size = area_country, colour = country_UN_continent) +
  #geom_text_repel(nudge_x = 0, nudge_y = 0, force = 0.03, force_pull = 2, seed = 123, min.segment.length = 0) +
  geom_text() +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") + 
  coord_trans(x = "log1p", y = "log1p") +
  scale_size_continuous(trans = "sqrt", guide = "none", range = c(3, 7)) +
  scale_colour_discrete(guide = "none") +
  scale_x_continuous(breaks = c(0, 10, 100, 1000, 1e4, 1e5, 1e6, 1e7), minor_breaks = NULL) +
  scale_y_continuous(breaks = c(0, 10, 100, 1000, 1e4, 1e5, 1e6, 1e7), minor_breaks = NULL) +
  theme_minimal() +
  labs(x = "Area PA according to clean column",
       y = "Area PA according to WDPA")
```

**Note**: We have large discrepancy for **Turkey**, so for now I will use `area_PA_WDPA`.

```{r corr PA_areas data}
data_rangers %>%
  filter(countryname_iso %in% c("TUR")) %>%
  select(contains("countryname"), "area_PA_WDPA", "area_PA_total")
```

**Question**: why is `area_PA_total` the same for both country? There seem to be an issue.

Let's look at the correlation between the number of rangers and     the PA area:

```{r corr area_PA_WDPA lm}
fit  <- lm(log(staff_rangers + 1) ~ log(area_PA_WDPA), data = data_rangers)
coef_fit <- coef(fit)
f <- function(x) x^coef_fit[2] * (exp(coef_fit[1]) - 1)
```

```{r corr area_PA_WDPA plot, fig.height=8, fig.width=8, cache=TRUE}
data_rangers %>%
  ggplot() +
  aes(y = staff_rangers, x = area_PA_WDPA, label = countryname, size = area_country, colour = country_UN_continent) +
  #geom_text_repel(nudge_x = 0, nudge_y = 0, force = 0.03, force_pull = 2, seed = 123, min.segment.length = 0) +
  geom_text() +
  coord_trans(x = "log", y = "log1p") +
  #stat_function(fun = f, colour = "blue", n = 1e5) +
  scale_x_continuous(breaks = c(10, 100, 1000, 1e4, 1e5, 1e6, 1e7), minor_breaks = NULL, limits = c(10, 1e7)) +
  scale_y_continuous(breaks = c(0, 10, 100, 1000, 1e4, 1e5, 1e6, 1e7), minor_breaks = NULL) +
  scale_size_continuous(trans = "sqrt", guide = "none", range = c(3, 7)) +
  scale_colour_discrete(guide = "none") +
  theme_minimal() +
  labs(x = "Area PA according to country",
       y = "Number of rangers")
```

**Conclusion**: there is a nice (expected) relationship between the number of rangers and the area of PA.
Since the slope of the linear model is `r round(coef_fit[2], 2)`, the number of rangers increases less quickly than the area of PA.

## area country

Let's first compare the area of the country and its PA:

```{r corr PA_country_areas plot, fig.height=8, fig.width=8, cache=TRUE}
data_rangers %>%
  ggplot() +
  aes(y = area_country, x = area_PA_WDPA, label = countryname, size = area_country, colour = country_UN_continent) +
  #geom_text_repel(nudge_x = 0, nudge_y = 0, force = 0.03, force_pull = 2, seed = 123, min.segment.length = 0) +
  geom_text() +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") + 
  coord_trans(x = "log1p", y = "log1p") +
  scale_x_continuous(breaks = c(0, 10, 100, 1000, 1e4, 1e5, 1e6, 1e7), minor_breaks = NULL) +
  scale_y_continuous(breaks = c(0, 10, 100, 1000, 1e4, 1e5, 1e6, 1e7), minor_breaks = NULL) +
  scale_size_continuous(trans = "sqrt", guide = "none", range = c(3, 7)) +
  scale_colour_discrete(guide = "none") +
  theme_minimal() +
  labs(x = "Area PA according to WDPA",
       y = "Area country")
```

Let's identify the countries with no PA according to WDPA:

```{r countries with no PA}
data_rangers %>%
  filter(area_PA_WDPA == 0) %>%
  select(countryname_eng, area_PA_WDPA, area_country) %>%
  arrange(desc(area_country))
```

**Question**: is the 0 valid for Somalia?


Let's look at the correlation between the number of rangers and the country area:

```{r corr area_country lm}
fit  <- lm(log(staff_rangers + 1) ~ log(area_country), data = data_rangers)
coef_fit <- coef(fit)
f <- function(x) x^coef_fit[2] * (exp(coef_fit[1]) - 1)
```

```{r corr area_country plot, fig.height=8, fig.width=8, cache=TRUE}
data_rangers %>%
  ggplot() +
  aes(y = staff_rangers, x = area_country, label = countryname, size = area_country, colour = country_UN_continent) +
  geom_text_repel(nudge_x = 0, nudge_y = 0, force = 0.03, force_pull = 2, seed = 123, min.segment.length = 0) +
  coord_trans(x = "log", y = "log1p") +
  #stat_function(fun = f, colour = "blue", n = 1e5) +
  scale_x_continuous(breaks = c(10, 100, 1000, 1e4, 1e5, 1e6, 1e7), minor_breaks = NULL, limits = c(10, 1e7)) +
  scale_y_continuous(breaks = c(0, 10, 100, 1000, 1e4, 1e5, 1e6, 1e7), minor_breaks = NULL) +
  scale_size_continuous(trans = "sqrt", guide = "none", range = c(3, 7)) +
  scale_colour_discrete(guide = "none") +
  theme_minimal() +
  labs(x = "Country area",
       y = "Number of rangers")
```

**Conclusion**: there is a nice (expected) relationship between the number of rangers and the area of the country.
Since the slope of the linear model is `r round(coef_fit[2], 2)`, the number of rangers increases less quickly than the country area.

Let's look at the correlation between the number of rangers and the proportion of country in PA :

```{r corr PA_prop plot, fig.height=8, fig.width=8, cache=TRUE}
data_rangers %>%
  filter(area_country > 1) %>%
    ggplot() +
    aes(y = staff_rangers, x = area_PA_WDPA/area_country, label = countryname, size = area_country, colour = country_UN_continent) +
    geom_text() +
    coord_trans(y = "log1p") +
    scale_y_continuous(breaks = c(0, 10, 100, 1000, 1e4, 1e5, 1e6, 1e7), minor_breaks = NULL) +
    scale_size_continuous(trans = "sqrt", guide = "none", range = c(3, 7)) +
    scale_colour_discrete(guide = "none") +
    theme_minimal() +
    labs(x = "Proportion of country in PA",
         y = "Number of rangers",
         size = "Country area")
```

**Conclusion**: no clear relationship.

## reliability

```{r corr reliability plot, fig.height=8, fig.width=8, cache=TRUE}
data_rangers %>%
    ggplot() +
    aes(y = staff_rangers, x = reliability, label = countryname, size = area_country, colour = country_UN_continent) +
    #geom_text_repel(seed = 123, min.segment.length = 0) +
    geom_text() +
    coord_trans(y = "log1p") +
    scale_y_continuous(breaks = c(0, 10, 100, 1000, 1e4, 1e5, 1e6, 1e7), minor_breaks = NULL) +
    scale_size_continuous(trans = "sqrt", guide = "none", range = c(3, 7)) +
    scale_colour_discrete(guide = "none") +
    theme_minimal() +
    labs(x = "Reliability score",
         y = "Number of rangers",
         colour = "continent")
```

**Conclusion**: no clear relationship. That's good!

## forest area pct

```{r corr forest_area plot, fig.height=8, fig.width=8, cache=TRUE}
data_rangers %>%
  ggplot() +
  aes(y = staff_rangers, x = area_forest_pct, label = countryname, size = area_country, colour = country_UN_continent) +
  geom_text() +
  coord_trans(y = "log1p") +
  scale_y_continuous(breaks = c(0, 10, 100, 1000, 1e4, 1e5, 1e6, 1e7), minor_breaks = NULL) +
  scale_size_continuous(trans = "sqrt", guide = "none", range = c(3, 7)) +
  scale_colour_discrete(guide = "none") +
  theme_minimal() +
  labs(x = "Percentage of forest area",
       y = "Number of rangers",
         colour = "continent")
```

**Conclusion**: no clear relationship. 


## ecological indexes

We start by comparing the ecological indexes:

```{r EVI_SPI plot, fig.height=8, fig.width=8, cache=TRUE}
data_rangers %>%
  ggplot() +
  aes(y = EVI, x = SPI, label = countryname, size = area_country, colour = country_UN_continent) +
  #geom_text_repel(min.segment.length = 0, nudge_x = 0, nudge_y = 0, force = 0.03, force_pull = 2, seed = 123) +
  geom_text() +
  coord_fixed(xlim = c(0, 100), ylim = c(0, 100)) +
  scale_size_continuous(trans = "sqrt", guide = "none", range = c(3, 7)) +
  scale_colour_discrete(guide = "none") +
  theme_minimal()
```

```{r EVI_EPI plot, fig.height=8, fig.width=8, cache=TRUE}
data_rangers %>%
  ggplot() +
  aes(y = EVI, x = EPI_2020, label = countryname, size = area_country, colour = country_UN_continent) +
  #geom_text_repel(min.segment.length = 0, nudge_x = 0, nudge_y = 0, force = 0.03, force_pull = 2, seed = 123) +
  geom_text() +
  coord_fixed(xlim = c(20, 80), ylim = c(20, 80)) +
  scale_size_continuous(trans = "sqrt", guide = "none", range = c(3, 7)) +
  scale_colour_discrete(guide = "none") +
  theme_minimal()
```


```{r EPI_SPI plot, fig.height=8, fig.width=8, cache=TRUE}
data_rangers %>%
  ggplot() +
  aes(y = EPI_2020, x = SPI, label = countryname, size = area_country, colour = country_UN_continent) +
  #geom_text_repel(min.segment.length = 0, nudge_x = 0, nudge_y = 0, force = 0.03, force_pull = 2, seed = 123) +
  geom_text() +
  coord_fixed(xlim = c(0, 100), ylim = c(0, 100)) +
  scale_size_continuous(trans = "sqrt", guide = "none", range = c(3, 7)) +
  scale_colour_discrete(guide = "none") +
  theme_minimal()
```

We then plot the relationship with the number of rangers.

### EVI

```{r corr EVI_rangers plot, fig.height=8, fig.width=8, cache=TRUE}
data_rangers %>%
  ggplot() +
  aes(y = staff_rangers, x = EVI, label = countryname, size = area_country, colour = country_UN_continent) +
  geom_text() +
  scale_size_continuous(trans = "sqrt", guide = "none", range = c(3, 7)) +
  scale_colour_discrete(guide = "none") +
  coord_trans(y = "log1p") +
  scale_y_continuous(breaks = c(0, 10, 100, 1000, 1e4, 1e5, 1e6, 1e7), minor_breaks = NULL) +
  theme_minimal() +
  labs(y = "Number of rangers")
```

### SPI

```{r corr SPI_rangers plot, fig.height=8, fig.width=8, cache=TRUE}
data_rangers %>%
  ggplot() +
  aes(y = staff_rangers, x = SPI, label = countryname, size = area_country, colour = country_UN_continent) +
  geom_text() +
  scale_size_continuous(trans = "sqrt", guide = "none", range = c(3, 7)) +
  scale_colour_discrete(guide = "none") +
  coord_trans(y = "log1p") +
  scale_y_continuous(breaks = c(0, 10, 100, 1000, 1e4, 1e5, 1e6, 1e7), minor_breaks = NULL) +
  theme_minimal() +
  labs(y = "Number of rangers")
```

### EPI

```{r corr EPI_rangers plot, fig.height=8, fig.width=8, cache=TRUE}
data_rangers %>%
  ggplot() +
  aes(y = staff_rangers, x = EPI_2020, label = countryname, size = area_country, colour = country_UN_continent) +
  geom_text() +
  scale_size_continuous(trans = "sqrt", guide = "none", range = c(3, 7)) +
  scale_colour_discrete(guide = "none") +
  coord_trans(y = "log1p") +
  scale_y_continuous(breaks = c(0, 10, 100, 1000, 1e4, 1e5, 1e6, 1e7), minor_breaks = NULL) +
  theme_minimal() +
  labs(y = "Number of rangers")
```

## GDP

We start by comparing the different GDP metrics:

```{r GDPcapita_2019 plot, fig.height=8, fig.width=8, cache=TRUE}
data_rangers %>%
  ggplot() +
  aes(y = GDP_capita, x = GDP_2019, label = countryname, size = area_country, colour = country_UN_continent) +
  # geom_text_repel(min.segment.length = 0, nudge_x = 0, nudge_y = 0, force = 0.03, force_pull = 2, seed = 123) +
  geom_text() +
  coord_trans(y = "log", x = "log") +
  scale_size_continuous(trans = "sqrt", guide = "none", range = c(3, 7)) +
  scale_y_continuous(breaks = 10^(0:20), minor_breaks = NULL) +
  scale_x_continuous(breaks = 10^(0:20), minor_breaks = NULL) +
  scale_colour_discrete(guide = "none") +
  theme_minimal()
```

```{r GDPcapita_growth plot, fig.height=8, fig.width=8, cache=TRUE}
data_rangers %>%
  ggplot() +
  aes(y = GDP_capita, x = GDP_growth, label = countryname, size = area_country, colour = country_UN_continent) +
  # geom_text_repel(min.segment.length = 0, nudge_x = 0, nudge_y = 0, force = 0.03, force_pull = 2, seed = 123) +
  geom_text() +
  coord_trans(y = "log") +
  scale_size_continuous(trans = "sqrt", guide = "none", range = c(3, 7)) +
  scale_y_continuous(breaks = 10^(0:20), minor_breaks = NULL) +
  scale_colour_discrete(guide = "none") +
  theme_minimal()
```

```{r GDP_growth plot, fig.height=8, fig.width=8, cache=TRUE}
data_rangers %>%
  ggplot() +
  aes(y = GDP_2019, x = GDP_growth, label = countryname, size = area_country, colour = country_UN_continent) +
  # geom_text_repel(min.segment.length = 0, nudge_x = 0, nudge_y = 0, force = 0.03, force_pull = 2, seed = 123) +
  geom_text() +
  coord_trans(y = "log") +
  scale_size_continuous(trans = "sqrt", guide = "none", range = c(3, 7)) +
  scale_y_continuous(breaks = 10^(0:20), minor_breaks = NULL) +
  scale_colour_discrete(guide = "none") +
  theme_minimal()
```

We then plot the relationship with the number of rangers

### GDP per capita

```{r corr GDP_capita_rangers plot, fig.height=8, fig.width=8, cache=TRUE}
data_rangers %>%
  ggplot() +
  aes(y = staff_rangers, x = GDP_capita, label = countryname, size = area_country, colour = country_UN_continent) +
  geom_text() +
  scale_size_continuous(trans = "sqrt", guide = "none", range = c(3, 7)) +
  scale_colour_discrete(guide = "none") +
  coord_trans(y = "log1p", x = "log") +
  scale_y_continuous(breaks = c(0, 10, 100, 1000, 1e4, 1e5, 1e6, 1e7), minor_breaks = NULL) +
  scale_x_continuous(breaks = 10^(0:20), minor_breaks = NULL) +
  theme_minimal() +
  labs(y = "Number of rangers")
```

### GDP

```{r corr GDP_2019_rangers plot, fig.height=8, fig.width=8, cache=TRUE}
data_rangers %>%
  ggplot() +
  aes(y = staff_rangers, x = GDP_2019, label = countryname, size = area_country, colour = country_UN_continent) +
  geom_text() +
  scale_size_continuous(trans = "sqrt", guide = "none", range = c(3, 7)) +
  scale_colour_discrete(guide = "none") +
  coord_trans(y = "log1p", x = "log") +
  scale_y_continuous(breaks = c(0, 10, 100, 1000, 1e4, 1e5, 1e6, 1e7), minor_breaks = NULL) +
  scale_x_continuous(breaks = 10^(0:20), minor_breaks = NULL) +
  theme_minimal() +
  labs(y = "Number of rangers")
```

**Note**: there is a strong relationship.

### GDP growth

```{r corr GDP_growth_rangers plot, fig.height=8, fig.width=8, cache=TRUE}
data_rangers %>%
  ggplot() +
  aes(y = staff_rangers, x = GDP_growth, label = countryname, size = area_country, colour = country_UN_continent) +
  geom_text() +
  scale_size_continuous(trans = "sqrt", guide = "none", range = c(3, 7)) +
  scale_colour_discrete(guide = "none") +
  coord_trans(y = "log1p") +
  scale_y_continuous(breaks = c(0, 10, 100, 1000, 1e4, 1e5, 1e6, 1e7), minor_breaks = NULL) +
  theme_minimal() +
  labs(y = "Number of rangers")
```

## unemployment

```{r corr unemployment_rangers plot, fig.height=8, fig.width=8, cache=TRUE}
data_rangers %>%
  ggplot() +
  aes(y = staff_rangers, x = unemployment, label = countryname, size = area_country, colour = country_UN_continent) +
  geom_text() +
  scale_size_continuous(trans = "sqrt", guide = "none", range = c(3, 7)) +
  scale_colour_discrete(guide = "none") +
  coord_trans(y = "log1p") +
  scale_y_continuous(breaks = c(0, 10, 100, 1000, 1e4, 1e5, 1e6, 1e7), minor_breaks = NULL) +
  theme_minimal() +
  labs(y = "Number of rangers")
```

## proportion of rural inhabitants

```{r corr rural_rangers plot, fig.height=8, fig.width=8, cache=TRUE}
data_rangers %>%
  ggplot() +
  aes(y = staff_rangers, x = rural_pct, label = countryname, size = area_country, colour = country_UN_continent) +
  geom_text() +
  scale_size_continuous(trans = "sqrt", guide = "none", range = c(3, 7)) +
  scale_colour_discrete(guide = "none") +
  coord_trans(y = "log1p") +
  scale_y_continuous(breaks = c(0, 10, 100, 1000, 1e4, 1e5, 1e6, 1e7), minor_breaks = NULL) +
  theme_minimal() +
  labs(y = "Number of rangers")
```

## income

Let's first compare the area of the GDP per capita with the income:

```{r GDP_capita_income plot, fig.height=8, fig.width=8, cache=TRUE}
data_rangers %>%
  ggplot() +
  aes(y = income, x = GDP_capita, label = countryname, size = area_country, colour = country_UN_continent) +
  # geom_text_repel(min.segment.length = 0, nudge_x = 0, nudge_y = 0, force = 0.03, force_pull = 2, seed = 123) +
  geom_text() +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") + 
  coord_trans(y = "log", x = "log") +
  scale_size_continuous(trans = "sqrt", guide = "none", range = c(3, 7)) +
  scale_y_continuous(breaks = 10^(0:20), minor_breaks = NULL) +
  scale_x_continuous(breaks = 10^(0:20), minor_breaks = NULL) +
  scale_colour_discrete(guide = "none") +
  theme_minimal()
```

**Note**: `income` and `GDP_capita` are very similar:

We then plot the relationship with the number of rangers:

```{r corr income_rangers plot, fig.height=8, fig.width=8, cache=TRUE}
data_rangers %>%
  ggplot() +
  aes(y = staff_rangers, x = income, label = countryname, size = area_country, colour = country_UN_continent) +
  geom_text() +
  scale_size_continuous(trans = "sqrt", guide = "none", range = c(3, 7)) +
  scale_colour_discrete(guide = "none") +
  coord_trans(y = "log1p", x = "log") +
  scale_y_continuous(breaks = c(0, 10, 100, 1000, 1e4, 1e5, 1e6, 1e7), minor_breaks = NULL) +
  theme_minimal() +
  labs(y = "Number of rangers")
```


## population density

Let's first compare the area of the country and its density:

```{r corr popdensity_country_areas plot, fig.height=8, fig.width=8, cache=TRUE}
data_rangers %>%
  ggplot() +
  aes(y = pop_density, x = area_country, label = countryname, size = area_country, colour = country_UN_continent) +
  #geom_text_repel(nudge_x = 0, nudge_y = 0, force = 0.03, force_pull = 2, seed = 123, min.segment.length = 0) +
  geom_text() +
  coord_trans(x = "log1p", y = "log1p") +
  scale_x_continuous(breaks = c(0, 10, 100, 1000, 1e4, 1e5, 1e6, 1e7), minor_breaks = NULL) +
  scale_y_continuous(breaks = c(0, 10, 100, 1000, 1e4, 1e5, 1e6, 1e7), minor_breaks = NULL) +
  scale_size_continuous(trans = "sqrt", guide = "none", range = c(3, 7)) +
  scale_colour_discrete(guide = "none") +
  theme_minimal() +
  labs(y = "Population density", x = "Area of the country")
```

**Note**: we retrieve the expected negative correlation.

We then plot the relationship with the number of rangers:

```{r corr popdensity_rangers plot, fig.height=8, fig.width=8, cache=TRUE}
data_rangers %>%
  ggplot() +
  aes(y = staff_rangers, x = pop_density, label = countryname, size = area_country, colour = country_UN_continent) +
  geom_text() +
  scale_size_continuous(trans = "sqrt", guide = "none", range = c(3, 7)) +
  scale_colour_discrete(guide = "none") +
  coord_trans(y = "log1p", x = "log") +
  scale_y_continuous(breaks = c(0, 10, 100, 1000, 1e4, 1e5, 1e6, 1e7), minor_breaks = NULL) +
  scale_x_continuous(breaks = c(0, 10, 100, 1000, 1e4, 1e5, 1e6, 1e7), minor_breaks = NULL) +
  theme_minimal() +
  labs(y = "Number of rangers")
```

## IUCN categories

Let's first compare the 2 IUCN measurements:

```{r corr IUCN plot, fig.height=8, fig.width=8, cache=TRUE}
data_rangers %>%
  ggplot() +
  aes(y = IUCN_1_4_prop, x = IUCN_1_2_prop, label = countryname, size = area_country, colour = country_UN_continent) +
  #geom_text_repel(nudge_x = 0, nudge_y = 0, force = 0.03, force_pull = 2, seed = 123, min.segment.length = 0) +
  geom_text() +
  geom_abline(slope = 1, intercept = 0, linetype = "dashed") + 
  scale_size_continuous(trans = "sqrt", guide = "none", range = c(3, 7)) +
  scale_colour_discrete(guide = "none") +
  theme_minimal()
```

We then plot the relationship with the number of rangers:

```{r corr IUCN12_rangers plot, fig.height=8, fig.width=8, cache=TRUE}
data_rangers %>%
  ggplot() +
  aes(y = staff_rangers, x = IUCN_1_2_prop, label = countryname, size = area_country, colour = country_UN_continent) +
  geom_text() +
  scale_size_continuous(trans = "sqrt", guide = "none", range = c(3, 7)) +
  scale_colour_discrete(guide = "none") +
  coord_trans(y = "log1p") +
  scale_y_continuous(breaks = c(0, 10, 100, 1000, 1e4, 1e5, 1e6, 1e7), minor_breaks = NULL) +
  theme_minimal() +
  labs(y = "Number of rangers")
```

```{r corr IUCN14_rangers plot, fig.height=8, fig.width=8, cache=TRUE}
data_rangers %>%
  ggplot() +
  aes(y = staff_rangers, x = IUCN_1_4_prop, label = countryname, size = area_country, colour = country_UN_continent) +
  geom_text() +
  scale_size_continuous(trans = "sqrt", guide = "none", range = c(3, 7)) +
  scale_colour_discrete(guide = "none") +
  coord_trans(y = "log1p") +
  scale_y_continuous(breaks = c(0, 10, 100, 1000, 1e4, 1e5, 1e6, 1e7), minor_breaks = NULL) +
  theme_minimal() +
  labs(y = "Number of rangers")
```
